Analyzing DECREASE trials for extent of data fabrication

### 2017-02-01 submission/manuscript.Rmd
Interim commit

---
title: "Analyzing DECREASE trials for extent of data fabrication"
author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    keep_tex: yes
  word_document: default
csl: ../bibliography/ama.csl
bibliography: ../bibliography/library.bib
---

@@

# pdf(sprintf('%sfigures/fig3.pdf', x12), width = 8, height = 6)
# 
# par(mar = c(2.5, 5, 0.5, 0.5),
#     mfrow = c(2, 2))
# plot(res_i_uncertain,
#      xlab = '',
#      ylab = 'Proportion',
#      ylim = 0:1,
#      type = 'l',
#      lty = 1)
# lines(res_i_certain, pch = 2, type = 'l', lty = 2)
# abline(h = c(.025, .975), lty = 3)
# 
# legend(x = 0, y = .9, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
# 
# plot(x = NA, y = NA,
#      xlim = c(0,1), ylim = c(0, length(res_i_certain)),
#      xlab = "Confidence in %", ylab = "Data points fabricated")
# 
# lines(x = seq(0, 1, .01), y = res_i_uncertain_lb)
# lines(x = seq(0, 1, .01), y = res_i_uncertain_ub)
# lines(x = seq(0, 1, .01), y = res_i_certain_lb, lty = 2, pch = 2, type = 'l')
# lines(x = seq(0, 1, .01), y = res_i_certain_ub, lty = 2, pch = 2, type = 'l')
# abline(v = .95, lty = 3)
# 
# # clip(x1 = 0, x2 = length(res_i_uncertain), y1 = 0.025, y2 = .975)
# # abline(v = which(abs(res_i_uncertain - .975) == min(abs(res_i_uncertain - .975))), lty = 1)
# # abline(v = which(abs(res_i_certain - .975) == min(abs(res_i_certain - .975))), lty = 2)
# 
# par(mar = c(4, 5, 0.5, 0.5))
# plot(res_iv_uncertain,
#      xlab = 'Data-points fabricated',
#      ylab = 'Proportion',
#      ylim = 0:1,
#      type = 'l',
#      lty = 1)
# lines(res_iv_certain, pch = 2, type = 'l',
#       lty = 2)
# abline(h = c(.025, .975), lty = 3)
# 
# plot(x = NA, y = NA,
#      xlim = c(0,1), ylim = c(0, length(res_iv_certain)),
#      xlab = "Confidence in %", ylab = "Data points fabricated")
# 
# lines(x = seq(0, 1, .01), y = res_iv_uncertain_lb)
# lines(x = seq(0, 1, .01), y = res_iv_uncertain_ub)
# lines(x = seq(0, 1, .01), y = res_iv_certain_lb, lty = 2, pch = 2, type = 'l')
# lines(x = seq(0, 1, .01), y = res_iv_certain_ub, lty = 2, pch = 2, type = 'l')
# abline(v = .95, lty = 3)
# invisible(dev.off())

@@

par(mar = c(5, 5, 2.5, 0),
    mfrow = c(2, 2))
plot(res_i_uncertain,
     xlab = 'Data points fabricated',
     ylab = 'Proportion',
     main = 'DECREASE I',
     ylim = 0:1,
     type = 'l',
     lty = 1)
lines(res_i_certain, pch = 2, type = 'l', lty = 2)
abline(h = c(.025, .975), lty = 3)

@@

plot(res_iv_uncertain,
     xlab = 'Data-points fabricated',
     ylab = '',
     main = 'DECREASE IV',
     ylim = 0:1,
     yaxt = 'n',
     type = 'l',
     lty = 1)
lines(res_iv_certain, pch = 2, type = 'l',
      lty = 2)
abline(h = c(.025, .975), lty = 3)

@@

```{r figure 3, fig.cap = 'Test', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
knitr::include_graphics(sprintf("%sfigures/fig3.pdf", x12))
```

```{r}
res_i_certain
res_i_uncertain
res_iv_certain
res_iv_uncertain
```

@@

@@

However, due to various types of beta-blockers, evidence might be moderated[@klei2015].

>>>>

---
title: "Analyzing DECREASE trials for extent of data fabrication"
author: "Chris HJ Hartgerink, Gerben ter Riet, Marleen Kemper"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
  word_document: default
csl: ../bibliography/ama.csl
bibliography: ../bibliography/library.bib
---

@@

# Old plotting. Saved temporarily. Ignore if forgotten to delete.
# pdf(sprintf('%sfigures/fig3.pdf', x12), width = 8, height = 6)
# 
# par(mar = c(2.5, 5, 0.5, 0.5),
#     mfrow = c(2, 2))
# plot(res_i_uncertain,
#      xlab = '',
#      ylab = 'Proportion',
#      ylim = 0:1,
#      type = 'l',
#      lty = 1)
# lines(res_i_certain, pch = 2, type = 'l', lty = 2)
# abline(h = c(.025, .975), lty = 3)
# 
# legend(x = 0, y = .9, bty = 'n', lty = c(1, 2), legend = c("Distribution estimate", "Point estimate"))
# 
# plot(x = NA, y = NA,
#      xlim = c(0,1), ylim = c(0, length(res_i_certain)),
#      xlab = "Confidence in %", ylab = "Data points fabricated")
# 
# lines(x = seq(0, 1, .01), y = res_i_uncertain_lb)
# lines(x = seq(0, 1, .01), y = res_i_uncertain_ub)
# lines(x = seq(0, 1, .01), y = res_i_certain_lb, lty = 2, pch = 2, type = 'l')
# lines(x = seq(0, 1, .01), y = res_i_certain_ub, lty = 2, pch = 2, type = 'l')
# abline(v = .95, lty = 3)
# 
# # clip(x1 = 0, x2 = length(res_i_uncertain), y1 = 0.025, y2 = .975)
# # abline(v = which(abs(res_i_uncertain - .975) == min(abs(res_i_uncertain - .975))), lty = 1)
# # abline(v = which(abs(res_i_certain - .975) == min(abs(res_i_certain - .975))), lty = 2)
# 
# par(mar = c(4, 5, 0.5, 0.5))
# plot(res_iv_uncertain,
#      xlab = 'Data-points fabricated',
#      ylab = 'Proportion',
#      ylim = 0:1,
#      type = 'l',
#      lty = 1)
# lines(res_iv_certain, pch = 2, type = 'l',
#       lty = 2)
# abline(h = c(.025, .975), lty = 3)
# 
# plot(x = NA, y = NA,
#      xlim = c(0,1), ylim = c(0, length(res_iv_certain)),
#      xlab = "Confidence in %", ylab = "Data points fabricated")
# 
# lines(x = seq(0, 1, .01), y = res_iv_uncertain_lb)
# lines(x = seq(0, 1, .01), y = res_iv_uncertain_ub)
# lines(x = seq(0, 1, .01), y = res_iv_certain_lb, lty = 2, pch = 2, type = 'l')
# lines(x = seq(0, 1, .01), y = res_iv_certain_ub, lty = 2, pch = 2, type = 'l')
# abline(v = .95, lty = 3)
# invisible(dev.off())

@@

par(mar = c(5, 5, 2.5, 0),
    mfrow = c(2, 2))
plot(res_i_uncertain,
     xlab = 'Data points fabricated',
     ylab = 'Proportion',
     main = 'DECREASE-I',
     ylim = 0:1,
     type = 'l',
     lty = 1)
lines(res_i_certain, pch = 2, type = 'l', lty = 2)
abline(h = c(.025, .975), lty = 3)

@@

plot(res_iv_uncertain,
     xlab = 'Data-points fabricated',
     ylab = '',
     main = 'DECREASE-IV',
     ylim = 0:1,
     yaxt = 'n',
     type = 'l',
     lty = 1)
lines(res_iv_certain, pch = 2, type = 'l',
      lty = 2)
abline(h = c(.025, .975), lty = 3)

@@

```{r figure 3, fig.cap = 'Inversion method results used to estimate the number of data points fabricated in the DECREASE-I and DECREASE-IV trials. The top row panels indicate $p_F$ (y-axis) for all $X$ out of $N$ fabricated data points (x-axis). The bottom row indicates the estimated number of fabricated data points (y-axis) when varying the degree of confidence (x-axis). Dotted lines indicate the bounds for a 95 percent CI.', out.width = '80%', fig.asp = .5, fig.align = 'center', echo = FALSE}
knitr::include_graphics(sprintf("%sfigures/fig3.pdf", x12))
```

For DECREASE-I, the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_i_certain - .025) - 1` - `r which.max(res_i_certain - .975) - 1`] or [`r which.min(res_i_uncertain - .025) - 1` - `r which.max(res_i_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively. The left column of Figure 3 depicts the $p_F$ per *X* fabricated data points (top panel) and the bounds of the confidence interval when the degree of confidence is altered (lower panel). Staying clearly between the dotted lines in the top panel, depicting the 95\% CI (top: .975; bottom: .025), it becomes apparent that the degree of uncertainty is too high to make any estimates with a high degree of confidence. This due to the small sample size (i.e., $N=`r length(res_i_certain) - 1`$) combined with the availability of just the summary results. Only when the degree of confidence is lowered to around 75\% does the interval not span the entire sample size. As such, based on the summary results, little can be said about the extent of the data fabrication that occurred in the DECREASE-I trial, affirming the conclusions of the original committee report [@commissie2013].

For DECREASE-IV, the 95\% confidence interval for the estimated number of fabricated data points is [`r which.min(res_iv_certain - .025) - 1` - `r which.max(res_iv_certain - .975) - 1`] or [`r which.min(res_iv_uncertain - .025) - 1` - `r which.max(res_iv_uncertain - .975) - 1`] when based on a point estimate or a more uncertain distribution estimate, respectively. The minor difference between the estimates indicates that there is a high degree of confidence that data fabrication did occur. Nonetheless, the range of potentially fabricated data points is still estimated at approximately 1000; this indicates that the summary results are insufficient to provide more than an estimated lowerbound. However, this does indicate that it is possible not all data were fabricated (i.e., $N=`r length(res_iv_certain) - 1`$), increasing the importance of data provenance to discern between genuine and falsified data. This affirms the conclusion of the scientific integrity committee that the results of the DECREASE-IV trial are "scientifically incorrect" (p.11) [@commissie2012].

The results also highlight that, despite the lack of availability of raw data, summary results from larger samples result in more certainty about the estimated number of fabricated data points using the inversion method. This is due to decreased standard errors at the population level of the effects, providing less variation in the simulations and resulting in higher sensitivity. However, this also highlights that in order to prevent detection it would be in the fabricators' interest to fabricate small studies (assuming the fabricator wants to remain undetected).

@@

Throughout steps 1, 2, and 3 we provided evidence that the , the DECREASE trials are highly unlikely to come from a genuine data pool, and that there is a high degree of confidence that data was fab

@@

However, due to various types of beta-blockers, evidence might be moderated [@klei2015].
